Apparent preferences for cognitive effort fade when multiple forms of effort and delay are interleaved in a foraging environment

### 2020-04-01 Cost_manuscript.Rmd
darft of exp 1 results

---
title: Apparent preferences for cognitive effort fade when multiple forms of effort and delay are interleaved in a foraging environment
author: "Claudio Toro-Serey^1^2 & Joseph T. McGuire^1^2"
csl: apa.csl
output:
  word_document: default #reference_docx: Style_reference_draft.docx
  pdf_document:
    highlight: haddock
    keep_tex: no
header-includes:
- \usepackage{setspace}\doublespacing
- \usepackage[left]{lineno}
- \linenumbers
- \usepackage{float}
- \newcommand{\beginsupplement}{ \setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}}
  \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}} }
#bibliography: mPFC_paper.bib
---

@@

```{r Setup, include = FALSE}
## libraries
library(tidyverse)
library(ggfortify)
library(knitr)
library(pander) 
library(nloptr)
library(pwr)
library(lme4)
#library(lmerTest)
library(gridExtra)
library(reshape2)
library(corrplot)

@@

## functions
# get the number of continuous quits before current trial
seqQuits <- function(subjData = 1){

@@

  # Prep list of results to be returned
  out <- list()
  out$percentQuit <- mean(choice == 0) * 100
  out$percentAccept <- mean(choice== 1) * 100 
  miss <- (choice != 1) & (choice != 0)
  out$percentMiss <- mean(miss)  * 100
  choice <- choice[!miss]
  reward <- as.numeric(reward[!miss])
  handling <- as.numeric(handling[!miss])

@@

  # Begin defining parameters
  # If choices are one-sided (i.e. all accepted)
  if ((sum(choice) == length(choice)) | (sum(choice) == 0)) {
    ifelse(sum(choice) == length(choice), out$Gamma <- exp(LB), out$Gamma <- exp(UB)) 
    out$Scale <- 1 # it was NA, but in theory a temperature of 1 also indicates noiseless estimates, and allows for easier fit computations
    out$LL <- 0
  } else {
    # Create a feasible region (search space)
    params <- as.matrix(expand.grid(scale = c(-1, 1), gamma = seq(LB, UB, length = 3)))
    # Create a list to check the minimization of the negative log lik.
    info <- list()
    info$negLL <- Inf
    # Define the options to be used during optimization
    # Consider looking into other optimization algorithms and global minima
    opts <- list("algorithm" = Algorithm,
               "xtol_rel" = 1.0e-8)
    # Optimize the sOC over all possible combinations of starting points
    for (i in seq(nrow(params))) {
      tempInfo <- nloptr(x0 = params[i, ], 
                   eval_f = optfun, 
                   lb = c(log(0.001), LB),
                   ub = c(-log(0.001), UB),
                   opts = opts,
                   choice = choice, 
                   handling = handling,
                   reward = reward)
      if (tempInfo$objective < info$negLL) {
        #print("Minimized")
        info$negLL <- tempInfo$objective
        info$params <- tempInfo$solution
      }
    }
    out$Gamma <- exp(info$params[2])
    out$Scale <- exp(info$params[1])
    out$LL <- -info$negLL
  }

  # Summarize the outputs
  out$LL0 <- log(0.5) * length(choice)
  out$Rsquared <- 1 - (out$LL/out$LL0)
  out$subjOC <- out$Gamma * handling
  out$p <- 1 / (1 + exp(-(out$Scale*(reward - out$subjOC))))
  out$predicted <- reward > out$subjOC
  out$predicted[out$predicted == TRUE] <- 1
  out$percentPredicted <- mean(out$predicted == choice)

  # adjust the probabilities in case of extreme gammas
  if (out$Gamma <= exp(LB)) {
    out$Gamma <- exp(LB) # temporary condition because Nlopt is not respecting the lower bound
    out$p <- rep(1, length(choice))
  } else if (out$Gamma == exp(UB)) {
    out$p <- rep(0, length(choice))
  }

@@

Cognitive and physical effort are typically regarded as costly, but recent findings have suggested that exerting effort can boost the value of prospects under certain conditions. Here we embedded mental and physical effort in a "diet choice" foraging task, which required decision makers not only to evaluate the magnitude and delay of a focal prospective reward, but also to estimate the general opportunity cost of time. In two experiments, independent sets of participants collected rewards that required equivalent periods of cognitive effort, physical effort, or unfilled delay. Monetary offers varied per trial, and the two experiments differed in whether the type of effort or delay cost was the same on every trial (between-subjects, n=21 per condition), or varied across trials (within-subjects, n=48). Participants were free either to accept or reject the cost/reward prospect offered on each trial. All participants were more likely to accept offers when rewards were higher and delays shorter, in line with a reward-maximizing strategy. Participants almost never reversed their acceptance decisions, and error rates in completing the effort requirement were low and decreased over time. When participants faced only one type of cost, cognitive effort persistently produced the highest acceptance rate compared to trials with an equivalent period of either physical effort or unfilled delay. This could be because cognitive effort was directly rewarding, or because it modulated foraging-related factors such as the perceived duration of delays or the estimated richness of the environment. We theorized that if cognitive effort were intrinsically rewarding, we would observe the same pattern of preferences when participants foraged for varying cost types in addition to rewards. In the within-subject experiment, an initially higher acceptance rate for cognitive effort trials disappeared over time amid an overall decline in acceptance rates as participants gained experience with all three conditions. Our results extend the view that cognitive demands may reduce the discounting effect of delays, but also suggest that differences in cost can eventually fade if individuals actively experience alternative forms of demand. Rather than assigning intrinsic value to cognitive effort, our findings support the idea that a cognitive effort requirement might influence contextual factors such as subjective delay durations or the perceived opportunity cost of time. Such altered estimations can be recalibrated if multiple forms of demand are interleaved.

@@

@@

  Both between- and within-subject experiments were pregeristered with the Open Science Framework (https://osf.io/2rsgm/registrations). All experimental procedures were approved by the Boston University Institutional Review Board, and written consent was acquired for all participants. For the between-subject experiment, we recruited individuals until a desired number of 84 eligible participants was achieved (58 Female, median age = 21, range = 18 - 31; number excluded before reaching goal = XX). The sample size was determined by means of power analysis (ANOVA), using a significance level of 0.05, power of 0.8, an effect size of f = 0.45 (calculated from a pilot study), and three groups (one for each cost type). The resulting per-group sample was 20, which we increased to 21 in order to match three possible block orders. We added an extra group of 21 participants who experienced a minimally effortful condition in order to determine whether effort or pure engagement were driving our results.
	We excluded participant datasets based on four preregistered criteria: 1) Consent: If they withdrew their participation; 2) Inattentiveness: a catch trial was placed at the end of each experimental block, asking participants to press a key within 3 seconds (time requirement based on pilot study response times). A participant that failed two or more of these checks was excluded and replaced. 3) Improbable choice behavior: The task was structured so that one reward amount must always be accepted. A participant who quit every trial in at least one block was assumed not to have followed or understood task instructions, or to have disengaged from the task altogether. 4) Performance: In the cognitive effort condition, participants were forced to travel if they made 2 mistakes in a trial (see task procedures below). Any participant with more than 30% forced travels was excluded.

@@

	Our foraging task was configured such that over and under accepting were detrimental to total earnings. To confirm this, we fit a general linear model with constant, linear, and quadratic terms to estimate the correspondence between proportion accepted (independent variable) and total earnings (dependent variable). A significant quadratic coefficient thus would signal that the task statistics operated as expected. Next, to determine the optimality of each group’s decisions, performed two-sided one-sample t-tests to see if the proportion of acceptances for each time/reward combination was significantly different from the optimal rate (see Operationalization of Cost above). This resulted in 36 independent tests (3 rewards amounts, 3 handling/travel time combinations, and 4 groups), so we corrected for multiple comparisons using False Discovery Rate (FDR) **CHANGE TO COST3 APPROACH.**
  Next, we compared preferences among cost conditions. We first performed a one-way ANOVA on the proportion of trials accepted using group as a factor. We then compared the proportion completed among all 4 groups using non-parametric permutation contrasts (6 tests). The same approach was used for total earnings. This gave us an initial glimpse on the potential differences in cost among conditions. In order to further look at the effect of handling, offer, and cost type, we computed the probability of accepting a trial with a mixed-effects logistic regression using the lme4 package (CITE). Based on the task structure and our main question, our a priori model of interest included cost condition, handling time, and reward amount as fixed main effects, and participant ID as a random effect. Cost condition was modeled with three categorical terms, with the fourth condition as the reference condition. We ran three versions of the model with different reference conditions, in order to test all pairwise differences among the four cost conditions. As preregistered, we anticipated significant main effects (coefficients different than zero) of handling time, reward, and cost condition. In addition, we hypothesized that the probability of accepting would be significantly great in either effort condition compared to wait and easy groups. We then examined whether this a priori model outperformed both simpler and more complex models. We used both Akaike’s Information Criterion (AIC) and Bayesian Information Criterion (BIC) to determine the model that minimized the negative log-likelihood while penalizing the addition of parameters. The regression with each combination of predictors was fitted in the following order: 1) intercept only; 2) cost type only; 3) handling time only; 4) reward only; 5) condition, handling, and reward main effects (a priori model); 6) adding a handling-by-reward interaction; and 7) adding all three possible two-way interactions. We predicted that model 5 would have the lowest AIC.
  Finally, we modeled the subjective opportunity cost in each condition. We started by addressing the prediction that participants would display consistent choice patterns throughout the session (i.e. that the reward amounts they accepted in a given timing condition would be similar throughout the experiment). This prediction was made because participants had full knowledge of the environment. We computed each participant’s total proportion of acceptances pre- and post-midpoint for every block. For each cost condition separately, we will fit a linear model that predicts post-midpoint acceptance from before-midpoint rates. We will report the slopes and 95% confidence intervals (CI) for each cost group. CIs containing 1 will denote that participants in that group produced consistent choices (CHANGE TO COST3 APPROACH, PERMUTATION).
  Once we established choice stability, we proceeded to estimate the subjective OC of each type of demand (see Operationalization of Cost above). OC were computed as the product of a free parameter (gamma) and the handling time. We used a softmax function to model each participant’s probability of completing a trial based on the difference between the delayed reward’s magnitude and the estimated OC for each cost type

@@

  Both gamma and the temperature parameter of the softmax function were estimated independently for each subject, using a maximum likelihood (MLE) approach through the NLOPTR package (CITE). We cross-validated each subject’s OC value using the pre-midpoint data for estimation, and post-midpoint choices for testing. The estimates from the training sample were used to predict acceptances in the testing sample, and the mean percent correctly predicted was reported for each group. Finally, the OC estimates for each group were compared using an ANOVA with condition as a factor. This let us determine which cost type produced the highest discounting. (ADD ADDITIONAL MODELS FROM LITERATURE)

@@

@@

  To compare acceptance rates among cost conditions, we first computed the probability of accepting a trial with a mixed-effects logistic regression. The model included cost condition and reward amount as fixed main effects, and subject ID as a random effect. Cost condition was modeled with three categorical terms, with the fourth condition as the reference condition (two effort conditions, and two delay conditions corresponding to each effort type). We ran three versions of the model with different reference conditions, in order to test all relevant pairwise differences among the four cost conditions. Next, we compared this model to both simpler and more complex models using Akaike’s Information Criterion (AIC) and Bayesian Information Criterion (BIC). We added varied model complexity in the following order: 1) intercept only; 2) condition only; 3) handling time only; 4) offer amount only; 5) handling time and offer amount (added to the preregistered comparison); 6) condition, handling time, and reward main effects (a priori model from above); 7) adding a handling-by-reward interaction; and 8) all possible two-way interactions. We predicted that model 6 would have the last considerable decrease in the negative log-likelihood. Nested models with similar AIC were statistically compared using an analysis-of-deviance. The significance test was computed as the probability of the reduction in deviance, based on a chi-square distribution with degrees of freedom equal to the difference in the number of parameters between models.
  Next we probed the confidence and consistency with which participants made choices. We computed the proportion of quits that were performed during the choice window versus during the handling time. We interpreted as confidence if a participant made at least 80% of their choices during the offer window. We then computed each participant’s total proportion of acceptances on the first two and last two blocks, and compared them using a paired permutation analysis (5000 iterations).
  To estimate the subjective opportunity cost of each condition, we will use a logistic function to model each participant’s probability of completing a trial based on the difference between the delayed reward’s magnitude and the estimated opportunity cost (OC) for each cost type. OC will be computed as the product of a free parameter (gamma) and the handling time. Both gamma and the temperature parameter of the logistic function will be estimated at the subject level, independently for each subject and cost type. We cross-validated each subject’s OC value using the first four blocks (two cognitive, two physical) for estimation, and the last two blocks (one of each effort type) choices for testing. The estimates were used to predict acceptances in the testing sample, and the mean percent correctly predicted was reported for each cost type. The OC estimates for each group will be compared using a repeated measures ANOVA with cost as a factor. We will perform pairwise paired permutations as post-hoc tests. This will let us determine which cost type produced the highest discounting.

@@

## plots
# pAccepted
p1 <- ggplot(temp, aes(Cost, pAccept, fill = Cost)) + 
        geom_boxplot(show.legend = F) +
        scale_fill_manual(values = colsBtw) +
        ylim(c(0, 1)) +
        labs(y = "Proportion Accepted") +
        theme(panel.grid.major = element_blank(), 
            panel.grid.minor = element_blank(), 
            panel.background = element_blank(), 
            axis.line = element_line(colour = "black"),
            text = element_text(size = 16))

# earnings
p2 <- ggplot(temp, aes(Cost, Earnings, fill = Cost)) + 
        geom_boxplot(show.legend = F) +
        scale_fill_manual(values = colsBtw) +
        labs(y = "Total Earnings (dollars)") +
        theme(panel.grid.major = element_blank(), 
            panel.grid.minor = element_blank(), 
            panel.background = element_blank(), 
            axis.line = element_line(colour = "black"),
            text = element_text(size = 16))

grid.arrange(p1, p2, ncol = 2)
```

>>>>

---
title: Apparent preferences for cognitive effort fade when multiple forms of effort and delay are interleaved in a foraging environment
author: "Claudio Toro-Serey^1,2 & Joseph T. McGuire^1,2"
csl: apa.csl
output:
  word_document: default #reference_docx: Style_reference_draft.docx
  pdf_document:
    highlight: haddock
    keep_tex: no
header-includes:
- \usepackage{setspace}\doublespacing
- \usepackage[left]{lineno}
- \linenumbers
- \usepackage{float}
- \newcommand{\beginsupplement}{ \setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}}
  \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}} }
#bibliography: mPFC_paper.bib
---

@@

```{r Setups, include = FALSE}
## libraries
library(tidyverse)
library(ggfortify)
library(knitr)
library(pander) 
library(nloptr)
library(pwr)
library(lme4)
#library(lmerTest)
library(gridExtra)
library(reshape2)
library(corrplot)
library(survival)
library(ggfortify)

@@

## functions
# parameter recovery function
test_params <- function(params = list(Gamma = 1), model = expr(params$Gamma * Handling), subjData, handling = c(2, 10, 14), offer = c(4, 8, 20), plot = F) {

  # this just computes choices made by a given parameterized model
  # or compares a putative model with participant choices
  #
  # params: a list of parameters that can be called by the model (besides behavioral observations from subjData)
  # model: must use expr() and call for extra parameters from params if needed (see default)
  # subjData: optional data to be compared
  # handling / offer: in case new experimental parameters should be tested (probably won't use)
  # plot: T/F whether to plot the comparison between modeled and actual behavior
  #

  # if no subject data was provided, use the default options and only output the recovered data
  if (missing(subjData)) {

    # compute
    df <- expand.grid(Handling = handling, Offer = offer) %>% 
      mutate(OC = eval(model),
             Choice = ifelse(Offer > OC, 1, 0))

    # plot
    if (plot) {

      ggplot(data = df, aes(interaction(Offer, Handling), Choice, color = Gamma)) +
        geom_line(aes(group = interaction(Handling, Gamma)), size = 1) +
        geom_point(size = 3) +
        labs(x = "Offer.Handling", y = "Proportion Accepted") +
        facet_wrap(vars(Gamma), ncol = 2) +
        theme(legend.key = element_blank(),
              panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(),
              panel.background = element_blank(),
              axis.line = element_line(colour = "black"),
              text = element_text(size = 16))

    # otherwise give a comparison between recovered and observed choice for a subject's data + gamma
  } else {

    # compute
    df <- subjData %>% 
      mutate(OC = eval(model),
             Choice = ifelse(Offer > OC, 1, 0))

    # plot
    if (plot) {

      print(ggplot(data = df) +
        geom_line(aes(interaction(Offer, Handling), propAccept, group = Handling), size = 1, color = "darkgreen") +
        geom_line(aes(interaction(Offer, Handling), Choice, group = Handling), size = 1, color = "blue") +
        geom_point(aes(interaction(Offer, Handling), propAccept), size = 3, color = "darkgreen") +
        geom_point(aes(interaction(Offer, Handling), Choice), size = 3, color = "blue") +
        labs(x = "Offer.Handling", y = "Proportion Accepted") +
        #scale_color_manual()
        theme(legend.key = element_blank(),
              panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(),
              panel.background = element_blank(),
              axis.line = element_line(colour = "black"),
              text = element_text(size = 16)))
    }

  return(df)

# get the number of continuous quits before current trial
seqQuits <- function(subjData = 1){

@@

  # Prep list of results to be returned
  out <- list()
  out$percentQuit <- mean(choice == 0) * 100
  out$percentAccept <- mean(choice == 1) * 100 
  miss <- (choice != 1) & (choice != 0)
  out$percentMiss <- mean(miss)  * 100
  choice <- choice[!miss]
  reward <- as.numeric(reward[!miss])
  handling <- as.numeric(handling[!miss])

@@

  # Begin defining parameters
  # If choices are one-sided (i.e. all accepted), assign the upper or lower bound
  if ((sum(choice) == length(choice)) | (sum(choice) == 0)) {
    ifelse(sum(choice) == length(choice), out$Gamma <- exp(LB), out$Gamma <- exp(UB)) 
    out$temperature <- 1 # it was NA, but in theory a temperature of 1 also indicates noiseless estimates, and allows for easier fit computations
    out$LL <- 0
  } else {
    # Create a feasible region (search space)
    params <- as.matrix(expand.grid(temperature = c(-1, 1), gamma = seq(LB, UB, length = 3)))
    # Create a list to check the minimization of the negative log lik.
    info <- list()
    info$negLL <- Inf
    # Define the options to be used during optimization
    # Consider looking into other optimization algorithms and global minima
    opts <- list("algorithm" = Algorithm,
               "xtol_rel" = 1.0e-8)
    # Optimize the sOC over all possible combinations of starting points
    for (i in seq(nrow(params))) {
      tempInfo <- nloptr(x0 = params[i, ], 
                   eval_f = optfun, 
                   lb = c(log(0.001), LB),
                   ub = c(-log(0.001), UB),
                   opts = opts,
                   choice = choice, 
                   handling = handling,
                   reward = reward)
      if (tempInfo$objective < info$negLL) {
        #print("Minimized")
        info$negLL <- tempInfo$objective
        info$params <- tempInfo$solution
      }
    }
    out$Gamma <- exp(info$params[2])
    out$temperature <- exp(info$params[1])
    out$LL <- -info$negLL
  }

  # Summarize the outputs
  out$LL0 <- log(0.5) * length(choice)
  out$Rsquared <- 1 - (out$LL/out$LL0) # pseudo r-squred, quantifying the proportion of deviance reduction vs chance
  out$subjOC <- out$Gamma * handling
  out$probAccept <- 1 / (1 + exp(-(out$Scale*(reward - out$subjOC))))
  out$predicted <- reward > out$subjOC
  out$predicted[out$predicted == TRUE] <- 1
  out$percentPredicted <- mean(out$predicted == choice)

  # adjust the probabilities in case of extreme gammas
  if (out$Gamma <= exp(LB)) {
    out$Gamma <- exp(LB) # temporary condition because Nlopt is not respecting the lower bound
    out$probAccept <- rep(1, length(choice))
  } else if (out$Gamma == exp(UB)) {
    out$probAccept <- rep(0, length(choice))
  }

@@

  Cognitive and physical effort are typically regarded as costly, but recent findings have suggested that exerting effort can boost the value of prospects under certain conditions. Here we embedded mental and physical effort in a "diet choice" foraging task, which required decision makers not only to evaluate the magnitude and delay of a focal prospective reward, but also to estimate the general opportunity cost of time. In two experiments, independent sets of participants collected rewards that required equivalent periods of cognitive effort, physical effort, or unfilled delay. Monetary offers varied per trial, and the two experiments differed in whether the type of effort or delay cost was the same on every trial (between-subjects, n=21 per condition), or varied across trials (within-subjects, n=48). Participants were free either to accept or reject the cost/reward prospect offered on each trial. All participants were more likely to accept offers when rewards were higher and delays shorter (in line with a reward-maximizing strategy), almost never reversed their acceptance decisions, and committed few errors while completing the effort requirements. When participants faced only one type of cost, cognitive effort persistently produced the highest acceptance rate compared to trials with an equivalent period of either physical effort or unfilled delay. We theorized that if cognitive effort were intrinsically rewarding, we would observe the same pattern of preferences when participants foraged for varying cost types in addition to rewards. In the within-subject experiment, an initially higher acceptance rate for cognitive effort trials disappeared over time amid an overall decline in acceptance rates as participants gained experience with all three conditions. Our results extend the view that cognitive demands may reduce the discounting effect of delays, but also suggest that differences in cost can eventually fade if individuals actively experience alternative forms of demand. Rather than assigning intrinsic value to cognitive effort, our findings support the idea that a cognitive effort requirement might influence contextual factors such as subjective delay durations or the perceived opportunity cost of time. Such altered estimations can be recalibrated if multiple forms of demand are interleaved.

@@

Effort is costly, but sometimes it can add value. In thinking about this, studies have focused on TAFC. We have learned much about this (show data), but two issues: 1) many studies compare costs by comparing discounting functions, which is helpful, but 2) an often disregarded aspect of this is OC. Even in tafc, a choice for a sooner or easier item can be a reflection of wanting to skip the offer altogether. Foraging provides a natural way to probe this. Foraging is often divided into two: patch and prey, and they are often experienced in daily life. Patch leaving can be deciding when to leave a job (diminishing gains). Prey can be X, or nerflix (show gif).

So going back to our question of the cost and value of effort, maybe the distinction has something to do with the ways in which we can allocate time and effort (work on this, including how foraging can provide a reference point to evaluate the value-modulatory effects of demands, and maybe put it before foraging is explained?).

In this talk, I'll show you how we'veused this framework to understand the inherent (elaborate on inherent above) and relative costly nature of effort and delay. I'll focus on behavioral findings, and I'll give a nudge to modeling. Then I'll discuss how we plan to understand the brain networks that characterize this behavior, with a focus on individual differences.

@@

  Both between- and within-subject experiments were pregeristered with the Open Science Framework (https://osf.io/2rsgm/registrations). All experimental procedures were approved by the Boston University Institutional Review Board, and written consent was acquired for all participants. For the between-subject experiment, we recruited individuals until a desired number of 84 eligible participants was achieved (58 Female, median age = 21, range = 18 - 31; number excluded before reaching goal = XX). The sample size was determined by means of power analysis (ANOVA), using a significance level of 0.05, power of 0.8, an effect size of f = 0.45 (calculated from a pilot study), and three groups (one for each cost type). The resulting per-group sample was 20, which we increased to 21 in order to match three possible block orders. We added an extra group of 21 participants who experienced a minimally effortful condition in order to determine whether effort or pure engagement were driving our results.

	We excluded participant datasets based on four preregistered criteria: 1) Consent: If they withdrew their participation; 2) Inattentiveness: a catch trial was placed at the end of each experimental block, asking participants to press a key within 3 seconds (time requirement based on pilot study response times). A participant that failed two or more of these checks was excluded and replaced. 3) Improbable choice behavior: The task was structured so that one reward amount must always be accepted. A participant who quit every trial in at least one block was assumed not to have followed or understood task instructions, or to have disengaged from the task altogether. 4) Performance: In the cognitive effort condition, participants were forced to travel if they made 2 mistakes in a trial (see task procedures below). Any participant with more than 30% forced travels was excluded.

@@

	Our foraging task was configured such that over and under accepting were detrimental to total earnings. To confirm this, we fit a general linear model with constant, linear, and quadratic terms to estimate the correspondence between proportion accepted (independent variable) and total earnings (dependent variable). A significant quadratic coefficient thus would signal that the task statistics operated as expected. Next, to determine the optimality of each group’s decisions, performed two-sided one-sample t-tests to see if the proportion of acceptances for each time/reward combination was significantly different from the optimal rate (see Operationalization of Cost above). This resulted in 36 independent tests (3 rewards amounts, 3 handling/travel time combinations, and 4 groups), so we corrected for multiple comparisons using False Discovery Rate (FDR). **MAYBE CHANGE TO COST3 APPROACH?**

  Next, we compared preferences among cost conditions. We first performed a one-way ANOVA on the proportion of trials accepted using group as a factor. We then compared the proportion completed among all 4 groups using non-parametric permutation contrasts (6 tests). On each of 5000 permutation iterations, the group assignment was randomly shuffled without replacement, and the difference in mean acceptance rates across iterations created a putative null distribution. The unpermuted group mean difference was then evaluated against this permuted distribution. The same approach was used for total earnings. This gave us an initial glimpse on the potential differences in cost among conditions. In order to further look at the effect of handling, offer, and cost type, we computed the probability of accepting a trial with a mixed-effects logistic regression using the lme4 package (CITE). Based on the task structure and our main question, our a priori model of interest included cost condition, handling time, and reward amount as fixed main effects, and participant ID as a random effect. Cost condition was modeled with three categorical terms, with the fourth condition as the reference condition. We ran three versions of the model with different reference conditions, in order to test all pairwise differences among the four cost conditions. As preregistered, we anticipated significant main effects (coefficients different than zero) of handling time, reward, and cost condition. In addition, we hypothesized that the probability of accepting would be significantly great in either effort condition compared to wait and easy groups. We then examined whether this a priori model outperformed both simpler and more complex models. We used both Akaike’s Information Criterion (AIC) and Bayesian Information Criterion (BIC) to determine the model that minimized the negative log-likelihood while penalizing the addition of parameters. The regression with each combination of predictors was fitted in the following order: 1) intercept only; 2) condition only; 3) handling time only; 4) offer amount only; 5) handling time and offer amount (added to the preregistered comparison); 6) condition, handling time, and reward main effects (a priori model from above); 7) adding a handling-by-reward interaction; and 8) all possible two-way interactions. We predicted that model 5 would have the lowest AIC. Nested models with similar AIC were statistically compared using an analysis-of-deviance. The significance test was computed as the probability of the reduction in deviance, based on a chi-square distribution with degrees of freedom equal to the difference in the number of parameters between models.

  Finally, we modeled the subjective opportunity cost in each condition. We started by addressing the prediction that participants would display consistent choice patterns throughout the session (i.e. that the reward amounts they accepted in a given timing condition would be similar throughout the experiment). We computed each participant’s total proportion of acceptances pre- and post-midpoint for every block, and performed paired permutations (5000) on the mean proportion of acceptances. We then confirmed that the relative costs observed with from the whole experimental session were still present on each half of the experiment. The winning model from the mixed-effects logistic comparison was applied to each half separately. In each case the reference cost category was rotated in order to assess pairwise differences among all costs. We evaluated the consistency of the relative costs by visually comparing the patterns of significant differences between experimental halves.

  Once we established choice stability, we proceeded to estimate the subjective OC of each type of demand (see Operationalization of Cost above). OC was computed as the product of a free parameter (gamma) and the handling time. Gamma can be interpreted as the subjective value of each second spent in an environment. If a trial's reward rate (i.e. offer amount divided by the handling time) fell below gamma, a participant was better off finding more profitable alternatives. In other words, individuals with higher gammas required higher rewards in order to accept a given offer. We used a softmax function to model each participant’s probability of completing a trial based on the difference between the delayed reward’s magnitude and the estimated OC for each cost type

@@

  Both gamma and the temperature parameter of the softmax function were estimated independently for each subject, using a maximum likelihood (MLE) approach through the NLOPTR package (CITE). The gamma optimization search space was bound by extreme choice values. For example, someone with an extremely high OC would not accept even the most beneficial offers, and would reject a 20 cent offer when the handling time was 2 seconds (or 20/2). Conversely, a participant experiencing low opportunity costs would accept the non-beneficial offer of 4 cents for a 14 second delay (or 4/14). We cross-validated each subject’s OC value using the pre-midpoint data for estimation, and post-midpoint choices for testing. The estimates from the training sample were used to predict acceptances in the testing sample, and the mean percent correctly predicted was reported for each group. Finally, the OC estimates for each group were compared using an ANOVA with condition as a factor. This let us determine which cost type produced the highest discounting. **ADD ADDITIONAL MODELS FROM LITERATURE**

@@

### Analyseses

@@

  To compare acceptance rates among cost conditions, we first computed the probability of accepting a trial with a mixed-effects logistic regression. The model included cost condition and reward amount as fixed main effects, and subject ID as a random effect (handling time is not modeled, as participants experienced a single time combination). Cost condition was modeled with three categorical terms, with the fourth condition as the reference condition (two effort conditions, and two delay conditions corresponding to each effort type). We ran three versions of the model with different reference conditions, in order to test all relevant pairwise differences among the four cost conditions. We compared the a priori model to alternative parameterizations using AIC and BIC, varying model complexity in the following order: 1) intercept only; 2) condition only; 3) reward only; 4) condition and reward main effects (a priori model); and 5) adding a two-way interaction. We predicted that model 4 would have the last considerable decrease in the negative log-likelihood. Models with similar AIC were formally compared using the analysis-of-deviance approach from the between-subjects experiment.

  Next we probed the confidence and consistency with which participants made choices. We computed the proportion of quits that were performed during the choice window versus during the handling time. We interpreted as confidence if a participant made at least 80% of their choices during the offer window. We then computed each participant’s total proportion of acceptances on the first two and last two blocks, and compared them using a paired permutation analysis (5000 iterations).
  **adapt once we decide on the model candidates**To estimate the subjective opportunity cost of each condition, we used a logistic function to model each participant’s probability of accepting a trial based on the difference between the delayed reward’s magnitude and the estimated opportunity cost (OC) for each cost type. As before, OC was computed as the product of a free parameter (gamma) and the handling time. Both gamma and the temperature parameter of the logistic function were estimated at the subject level, independently for each subject and cost type. We cross-validated each subject’s OC value using the first four blocks (two cognitive, two physical) for estimation, and the last two blocks (one of each effort type) choices for testing. The estimates were used to predict acceptances in the testing sample, and the mean percent correctly predicted was reported for each cost type. The OC estimates for each group were compared using a repeated measures ANOVA with cost as a factor. We performed pairwise paired permutations as post-hoc tests.

@@

## plots
# pAccepted
p1 <- ggplot(temp, aes(Cost, pAccept, fill = Cost)) + 
        geom_boxplot(show.legend = F) +
        geom_jitter(width = 0.1, alpha = 0.3, show.legend = F) +
        scale_fill_manual(values = colsBtw) +
        ylim(c(0, 1)) +
        labs(y = "Proportion Accepted") +
        theme(panel.grid.major = element_blank(), 
            panel.grid.minor = element_blank(), 
            panel.background = element_blank(), 
            axis.line = element_line(colour = "black"),
            text = element_text(size = 16))

# earnings
p2 <- ggplot(temp, aes(Cost, Earnings, fill = Cost)) + 
        geom_boxplot(show.legend = F) +
        geom_jitter(width = 0.1, alpha = 0.3, show.legend = F) +
        scale_fill_manual(values = colsBtw) +
        labs(y = "Total Earnings (dollars)") +
        theme(panel.grid.major = element_blank(), 
            panel.grid.minor = element_blank(), 
            panel.background = element_blank(), 
            axis.line = element_line(colour = "black"),
            text = element_text(size = 16))

grid.arrange(p1, p2, ncol = 2)

# Proportion of performance-based forced travels per subject and cost type
# move to 16.2.1.
propFails <- dataBtw %>% 
  filter(Cost != "Wait") %>% 
  group_by(SubjID, Cost) %>% 
  summarise(propFails = mean(rawChoice == 2)) %>%
  group_by(Cost) %>% 
  summarise(meanFT = mean(propFails),
            sdFT = sd(propFails))

## MAYBE BETTER WAY TO DO PAIRWISE COMPARISONS, USING VECTORIZE FROM BELOW
# OCanova$perms <- summaryOC$all %>%
#   select(Cost, Gamma) %>%
#   plyr::dlply("Cost", identity) %>%
#   do.call(cbind, .) %>%
#   select_if(is.numeric) %>%
#   rename(Physical = Physical.Gamma,
#          Cognitive = Cognitive.Gamma,
#          Wait = Wait.Gamma,
#          Easy = Easy.Gamma) %>%
#   outer(., ., FUN = Vectorize(permute), simple = T)
# diag(OCanova$perms) <- 1

  The next set of analyses examined the hypothesis that choice differences among costs would be characterized by more acceptances by the cognitive group **not quite**. Figure X shows that the cognitive effort group consistently accepted more offers, which resulted in lower earnings. One-way ANOVAs showed that differences among groups were significant both for overall acceptance rates (F = `r round(propAov[[1]][1, 4], digits = 2)`, *p* `r report_p(propAov[[1]][1, 5])`), and total earnings (F = `r round(earnAov[[1]][1, 4], digits = 2)`, *p* `r report_p(earnAov[[1]][1, 5])`). Post-hoc permutations (5000 iterations) comparing mean acceptance rates between every pair of costs confirmed that acceptance rates of the cognitive effort group (mean = `r filter(overallSumms, Cost == "Cognitive") %>% select(mProp)`, SD = `r filter(overallSumms, Cost == "Cognitive") %>% select(sdProp)`) were higher than those of the physical (mean = `r filter(overallSumms, Cost == "Physical") %>% select(mProp)`, SD = `r filter(overallSumms, Cost == "Physical") %>% select(sdProp)`; permuted *p* = `r report_p(filter(propTests, contrast == "Physical < Cognitive") %>% select(permPval))`, Cohen's D = `r filter(propTests, contrast == "Physical < Cognitive") %>% select(cohenD)`) and wait groups (mean = `r filter(overallSumms, Cost == "Wait") %>% select(mProp)`, SD = `r filter(overallSumms, Cost == "Wait") %>% select(sdProp)`; permuted *p* = `r report_p(filter(propTests, contrast == "Cognitive > Wait") %>% select(permPval))`, Cohen's D = `r filter(propTests, contrast == "Cognitive > Wait") %>% select(cohenD)`). It also showed that those faced with the easy task accepted more than those in the wait group, although responses from the easy group were more variable (mean Easy = `r filter(overallSumms, Cost == "Easy") %>% select(mProp)`, SD = `r filter(overallSumms, Cost == "Easy") %>% select(sdProp)`; permuted *p* = `r report_p(filter(propTests, contrast == "Wait < Easy") %>% select(permPval))`, Cohen's D = `r filter(propTests, contrast == "Wait < Easy") %>% select(cohenD)` (all other *p* > 0.05). The proportion of forced travels for cognitive effort was `r round(propFails[2, 2], digits = 2)` across participants (and 0 for either physical or easy conditions), making performance an unlikely explanation for the observed differences in choice.
